The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-25 paper.tex
changes in anthropomorphism - artifact-centered section

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts / the design of the non-human agent.
The direction of causality in this view is that the non-human artifact (or a certain 
feature of it) encourages the human to anthropomorphize the artifact / agent.
This perspective assumes that humans directly respond to life-like and social cues 
that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it. In fact, from early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
and to respond to social cues, such as facial expressions, social gestures or turn-
taking. When first looking at an unknown object, the visual scope of its design, 
its outer appearance is likely to have an important impact on how the object is 
overall perceived (how it appears to the human) \citep{schmitz_concepts_2011}.

This fact is exploited in the design of robots and other things. The basic 
assumption  here is that if an artifact 
somehow resembles a human, it is likely to be treated similar to a human. If this 
explanation of anthropomorphism is correct, people may respond automatically to 
social cues emitted by an artificial embodied agent such as a robot, and apply 
human-human social schemas and norms to the occurring interactions. \hl{(However, at 
the same time, the human-like design of things also raises expectations from the 
human user side, to which we will come back later in more detail.)}

@@

The use of anthropomorphic (human-like) form in the design of non-human 
things is as old as people's tendency to anthropomorphize. Anthropomorphic (and 
also animal-shaped) forms were used already in prehistoric artworks and 
illustrations, and they have a long-standing tradition in the design of daily life 
artifacts and in general \citep{guthrie_bottles_2000}. The principle of human-like 
form is probably most prominent in the design of robots. But what do we mean by \textit{human-like / anthropomorphic form}?

Using the term \textit{``form''} to describe the design of an artifact, we refer 
not only to the shape of an artifact (how the artifact looks like) but to the total 
expression of the product \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. Latter aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}. 
However, a classification makes sense because different user reactions are 
expected. According to \cite{bartneck_shaping_2004}, the appearance and function of 
a product, whether it is an appliance or a humanoid robot, affects the way that 
people perceive it, interact with it, and build long-term relationships with it. The relation between the design of a robot and the user's perception of it, is also formulated in the \textit{Uncanny Valley} hypothesis \citep{mori_uncanny_1970}.

@@

In general, it is suggested that increasing human-likeness of a robot will lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}). Humans seem to readily accept and prefer the 
anthropomorphic form of a robot, when compared to purely mechanical robots. A 
possible explanation for this is that 
humans tend to apply the same communication references to anthropomorphic systems 
that they would use with other humans. This theoretical concept is also supported 
by psychological determinants of 
perceiving human-likeness in something non-human: \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. 
Also various 
empirical studies have shown that robots with human-like forms can have a positive 
effect, as they enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

@@

\subsubsection{Anthropomorphism in Robotics and Human-Robot Interaction\\}
\label{sec:anthropomorphism-robotics}

\hl{Julia: I need to work on this section ...}

%While the previous sections presented anthropomorphism more generally, this 
%section focuses on how anthropomorphism is understood and has been studied in human 
%robot interaction / social robotics. Our aim here is to give a rough overview of 
%the diverse and partially conflicting interpretations of anthropomorphism in HRI, 
%and critically review some previous work.

Most work that studies anthropomorphism in HRI builds on the famous work of \cite{reeves_media_1996} in HCI. In the so-called \textit{Media Equation}, the authors describe that people mindlessly react socially to their computer and that they tend to even attribute human (social) characteristics to it. Apart from computers, anthropomorphism has also been observed with digital / virtual characters \citep{reeves_media_1996,
nass_anthropocentrism_1995}, and as more interactive technologies and systems emerged, also with artificial embodied agents, such as robots.

In human-robot interaction anthropomorphism is a commonly observed phenomenon, however, findings have been inconsistent, and in turn, anthropomorphic designs of robots are controversially discussed.

For instance, recent studies found that people tend to ascribe personality and intentions to robots \hl{(REFS)} and that they show empathetic behavior toward robots \citep{rosenthal-vonderputten_experimental_2013}. Summing it up, there is a huge and constantly growing body of literature that describes anthropomorphism in HCI and HRI. Despite the phenomenon is well studied, we still know rather little about it, so that at conferences and symposiums, anthropomorphism is still a ``hot topic''; however, at the same time theories related to anthropomorphism, such as the \textit{``Uncanny Valley''} are controversially discussed.

@@

But what is the motivation which underlies the efforts of enhancing anthropomorphism through specific anthropomorphic design in robots? Why does anthropomorphism seem to be so important in robotics? There is no one single answer to this question but certainly, developers and designers have discovered that anthropomorphism seems to be beneficial for human-robot relations, especially with socially interactive robots \cite{fong_survey_2003}.

The domain of (social) robotics tries to exploit this fact by enhancing anthropomorphism through anthropomorphic robots: \textit{"Applying anthropomorphism or zoomorphism changes the way how users try to understand and make sense of a [system] by projecting everyday \hl{expectations} of human and animalistic life onto it."} \cite{schmitz_concepts_2011}.

Studies have further shown that perceived similarity between the user and the robot influences the intensity of the participants' attributions towards the robot. Thus, anthropomorphism plays an important role in the design of robots as it is strongly related to the perception of intelligence, fun, the attribution of intentions and thus the predictability of the robot \hl{which ref??}.

@@

\subsection{A human-centered perspective on anthropomorphism}

>>>>

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts / the design of the non-human agent.
The direction of causality in this view is that the non-human artifact (or a certain 
feature of it) encourages the human to anthropomorphize the artifact / agent.
This perspective assumes that humans directly respond to life-like and social cues 
that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it. Humans naturally react to specific 
cues. This is natural because, for instance, from early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-
taking. 
These cues can be of various senses and modalities. One of the most important aspects 
is the outer appearance of the unknown agent or artifact. \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived (how it appears to the human). Thus, if a robot looks much like a human, it appears to be like a human, and is perceived accordingly.

This fact is exploited in the design of robots and other things. Again, the basic 
assumption here is that if an artifact 
somehow resembles a human, it is likely to be treated similar to a human. If this 
explanation of anthropomorphism is correct, people may respond automatically also to 
social cues emitted by an artificial embodied agent such as a robot, and apply 
human-human social schemas and norms to the occurring interactions. \hl{(However, at 
the same time, the human-like design of things also raises expectations from the 
human user side, to which we will come back later in more detail.)}
The human-like design of robots implies various consequences, of which we will present some in the following.

@@

What do we mean by \textit{human-like / anthropomorphic form}?
Using the term \textit{``form''} to describe the design of an artifact, we refer 
not only to the shape of an artifact (how the artifact looks like) but to the total 
expression of the product \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. Latter aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.

The use of anthropomorphic (human-like) form in the design of non-human 
things is as old as people's tendency to anthropomorphize. Anthropomorphic (and 
also animal-shaped) forms were used already in prehistoric artworks and 
illustrations, and they have a long-standing tradition in the design of daily life 
artifacts and in general \citep{guthrie_bottles_2000}. The principle of human-like 
form is probably most prominent in the design of robots, due to several reasons. 
However, it has been controversially discussed since the early 1970ies, related to 
\cite{mori_uncanny_1970} theory of the \texit{Uncanny Valley}, see Figure \ref{fig:uncanny_valley}.

The underlying crucial point about the design are the implications that it holds, and the 
suggestions for interactions that it makes to the user. In other words, the appearance 
and function of a product, whether it is an appliance or a humanoid robot, affects the 
way that people perceive it, interact with it, and build long-term relationships with 
it \cite{bartneck_shaping_2004}. The relation between the design of a robot and the 
user's perception of it, is also formulated in the \textit{Uncanny Valley} hypothesis \citep{mori_uncanny_1970}.

@@

Despite the fact that the Uncanny Valley theory was more a working hypothesis and not 
meant to be empirically examined \citep{cheetham_human_2011}, it has been studied quite 
extensively. However, results are inconsistent. We try to give a short overview (not 
exclusive) on work related to HRI and robotics in the following.

\subsubsection{How people perceive and interact with human-like robots\\}
\label{sec:anthropomorphism-robotics}

Most work that studies how far human users treat robots as social agents, builds on the 
famous work of \cite{reeves_media_1996} in HCI. In the so-called \textit{Media 
Equation}, the authors describe that people mindlessly react in a social way to their computer 
and that they tend to even attribute human (social) characteristics to it -- despite 
the fact that a computer does not have human-like characteristics. Apart from 
computers, anthropomorphism has also been observed with digital / virtual characters 
\citep{reeves_media_1996,nass_anthropocentrism_1995}, and artificial embodied agents, 
such as robots. In short, one can say that with robots, and human-like robots in 
particular, anthropomorphism is a commonly observed phenomenon. However, the 
anthropomorphic design approach in robotics is controversially discussed.

But what are the motivations which underlie the efforts of enhancing anthropomorphism through specific anthropomorphic design in robots? Why are anthropomorphic forms so distinct and at the same time controversial in robotics? There is no one single answer to this question.

\hl{human-likeness has a positive effect\\}

On one hand, developers and designers have discovered that human-like robots may better match the affordances that our to humans adapted world brings along. Also. anthropomorphic form and people's resulting consequence to anthropomorphize, appears to be generally beneficial for human-robot relations, especially with socially interactive robots \cite{fong_survey_2003}. Consequently, the domain of (social) robotics tries to exploit this fact by enhancing anthropomorphism through anthropomorphic robots: \textit{"Applying anthropomorphism or zoomorphism changes the way how users try to understand and make sense of a [system] by projecting everyday \hl{expectations} of human and animalistic life onto it."} \cite{schmitz_concepts_2011}.

In general, it is suggested that increasing the human-likeness of a robot (up to a 
certain threshold) will lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}). Humans seem to readily accept and prefer the 
anthropomorphic form of a robot, when compared to purely mechanical robots. A 
possible explanation for this is that 
humans tend to apply the same communication references to anthropomorphic systems 
that they would use with other humans, and this is what they are familiar with and feel comfortable about. This theoretical concept is also supported 
by psychological determinants of 
perceiving human-likeness in something non-human: \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. (We will come back to this later, in Section \ref{sec:explanations}).

Studies have further shown that perceived similarity between the user and the robot influences the intensity of the participants' attributions towards the robot. Thus,the anthropomorphic design of a robot impacts the user's perception of the robot, and how far the user is likely to attribute and experience qualities such as intelligence, fun, or intentions in the robot.

Various 
empirical studies have shown that robots with human-like forms can have a positive 
effect, as they enhance social 
responses from humans which in turn can positively impact on acceptance 
\citep{duffy_anthropomorphism_2003,goetz_cooperation_2002,venkatesh_theoretical_2000}. 
People responded more positively to an artifact that displayed human-like 
behavioral characteristics (emotions, facial expression) in contrast to a purely 
functional design 
\citep{eyssel_anthropomorphic_2010,krach_can_2008,reeves_media_1996,riek_how_2009}. 
Overall, these results suggest that human-like features in robots enhance 
anthropomorphism, and in turn create a social view toward the robot.

\hl{human-likeness has a negative effect\\}

@@

%For instance, recent studies found that people tend to ascribe personality and 
%intentions to robots \hl{(REFS)} and that they show empathetic behavior toward robots 
%\citep{rosenthal-vonderputten_experimental_2013}. Summing it up, there is a huge and 
%constantly growing body of literature that describes anthropomorphism in HCI and HRI. 
%Despite the phenomenon is well studied, we still know rather little about it, so that 
%at conferences and symposiums, anthropomorphism is still a ``hot topic''; however, at 
%the same time theories related to anthropomorphism, such as the \textit{``Uncanny 
%Valley''} are controversially discussed.

@@

@@

\subsection{A human-centered perspective on anthropomorphism}
\label{sec:human-centered}
